An FPGA Implementation of a Long Short-Term Memory Neural Network

### 2016-10-09 paper/paper.tex
Conference version of paper

    \item \textbf{Block Input Gate} -- as the name implies, this gate controls the flow of information
from the input gate to the memory cell. It also receives the input vector and the previous output
set, producing an output $\mb{z}^{(t)}$. The \textbf{activation function of this gate can both} the logistic
sigmoid, $\sigma(x) = 1 \over (1+e^{-x})$, or the hyperbolic tangent, $\tanh(x)$ but the most common choice is the
\textbf{hyperbolic hangent}.

@@

Hitherto, there is but one actual implementation of an LSTM network in hardware, published recently
(March 2016) by Chang et al.~\cite{Chang15}. A 2 layer LSTM network (two layers in series) was implemented in FPGA,
with 128 neurons each, which, for processing 1000 samples, yielded an execution time of \SI{0.932}{\second}. Assuming
that both layers are equal, and thus have the same execution time, this yields an approximate execution time of
$29.13 \, \si{\micro\second}$ per incoming sample (dividing the total execution time by $2 \times 1000$).
Therefore, the network~\cite{Chang15} is able to perform around \num{34.3d3} forward propagations per second.
Although the network is large, its performance is well below this work, which achieves \num{487d3} forward propagations per second, for that network size.
This is because the work in~\cite{Chang15} does not have a full level of parallelism when compared to the proposed design of
Section~\ref{sec:proparch}, and it makes use of external memory to store the weights. On the other hand, the authors use it as a co-processor for the main CPU, and not as a
standalone implementation.

@@

\begin{equation}\label{eq:kg}
K_G = \frac{\text{Number of rows}}{\text{Number of multipliers}}.
\end{equation}
This parameters represents the number of rows that share the same multiplier. The
same multiplier is responsible for producing several $i$-th elements of the output vector, in consecutive time slots of $M$ clock cycles.
For instance, in an $8\times2$ matrix scenario, with $K_G = 2$, we would have 4 multipliers, and the output vector
elements $y_0$, $y_2$, $y_4$ and $y_6$ would be ready after $M=2$ clock cycles, and the remaining -- $y_1$, $y_3$, $y_5$ and $y_7$ --  are
ready after another two clock cycles, that is $2M = 4$ clock cycles after the calculations began.

@@

The learning problem presented to both the software and hardware network is the \textbf{addition} of two binary numbers of 8 bits. The $i$-th
bit of each number is fed to the network as a vector, and the network outputs its prediction of the correct value of the $i$-th bit
of the \emph{result}. After the whole number is processed, the memory cells of the LSTM network are reset and a new addition
task can be presented to the network. Even though this seems a rather simple problem, it accounts for all the essential issues at which this
network excels: first, this is a \emph{classification} problem in which the Machine Learning algorithm needs to output a prediction based on
the input feature vector, and that prediction has to take into account the \emph{history} of predictions and inputs so far, because the current
bit is the sum of not only the bits of the two operands, but also the \emph{carry} generated at the last few positions -- this is where the
memory cells of this special Recursive Neural Network come into effect.

@@

\begin{table}
	\caption{Maximum clock frequencies for the VC707}
	\label{tab:maxfreq-virtx7}
    \centering
  \begin{tabular}{ | l | c | }
    \hline
     & $Frequency$  \\
    \hline
    $N=64$ & 140.854 MHz \\
    \hline
    $N=128$ & 140.854 MHz\\
    \hline
\hline
  \end{tabular}

>>>>

    \item \textbf{Block Input Gate} -- as the name implies, this gate controls the flow of information
from the input gate to the memory cell. It also receives the input vector and the previous output
set, producing an output $\mb{z}^{(t)}$. The \textbf{activation function of this gate can both} the logistic
sigmoid, $\sigma(x) = 1 \over (1+e^{-x})$, or the hyperbolic tangent, $\tanh(x)$ but the most common choice is the
\textbf{hyperbolic tangent}.

@@

Hitherto, there is but one actual implementation of an LSTM network in hardware, published recently
(March 2016) by Chang et al.~\cite{Chang15}. A 2 layer LSTM network (two layers in series) was implemented in FPGA,
with 128 neurons each, which, for processing 1000 samples, yielded an execution time of \SI{0.932}{\second}. Assuming
that both layers are equal, this yields an approximate execution time of $466 \, \si{\micro\second}$ per incoming sample (dividing the total execution time by $2 \times 1000$), and
if the computation time increases linearly with time, an 8 neuron layer would have an execution time of $29.13 \, \si{\micro\second}$.
Therefore, an 8-neuron network of~\cite{Chang15} would be able to perform around \num{34.3d3} forward propagations per second, while this work achieves \num{487d3} forward propagations per second, for that network size. This is because the work in~\cite{Chang15} does not have a full level of parallelism when compared to the proposed design of
Section~\ref{sec:proparch}, and it makes use of external memory to store the weights. On the other hand, the authors use it as a co-processor for the main CPU, and not as a
standalone implementation.

@@

\begin{equation}\label{eq:kg}
K_G = \frac{\text{Number of rows}}{\text{Number of multipliers}}.
\end{equation}
This parameters represent the number of rows that share the same multiplier. The
same multiplier is responsible for producing several $i$-th elements of the output vector, in consecutive time slots of $M$ clock cycles.
For instance, in an $8\times2$ matrix scenario, with $K_G = 2$, we would have 4 multipliers, and the output vector
elements $y_0$, $y_2$, $y_4$ and $y_6$ would be ready after $M=2$ clock cycles, and the remaining -- $y_1$, $y_3$, $y_5$ and $y_7$ --  are
ready after another two clock cycles, that is $2M = 4$ clock cycles after the calculations began.

@@

The learning problem presented to both the software and hardware network is the \textbf{addition} of two binary numbers of 8 bits. The $i$-th
bit of each number is fed to the network as a vector, and the network outputs its prediction of the correct value of the $i$-th bit
of the \emph{result}. After the whole number is processed, the memory cells of the LSTM network are reset and a new addition
task can be presented to the network. Even though this seems a rather simple problem, it accounts for all the essential issues at which this
network excels: first, this is a \emph{classification} problem in which the Machine Learning algorithm needs to output a prediction based on
the input feature vector, and that prediction has to take into account the \emph{history} of predictions and inputs so far, because the current
bit is the sum of not only the bits of the two operands, but also the \emph{carry} generated at the last few positions -- this is where the
memory cells of this special Recurrent Neural Network come into effect.

@@

\begin{table}
	\caption{Maximum clock frequencies for the VC707}
	\label{tab:maxfreq-virtx7}
    \centering
  \begin{tabular}{ | l | c | }
    \hline
     & $Frequency$  \\
    \hline
    $N=64$ & 140.854 MHz \\
    \hline
    $N=128$ & 140.854 MHz\\
    \hline
  \end{tabular}
